---
title: "Analysis of BMI GIANT GWAS data"
author: "Simina Boca, Jeff Leek"
date: "`r Sys.Date()`"
output: rmarkdown::pdf_document
vignette: >
  %\VignetteIndexEntry{swfdr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!--
%% \VignetteEngine{knitr::rmarkdown}
-->

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.path='Figures/')
```

```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown()
```

Load the relevant libraries:

```{r,message=FALSE}
library(readr)
library(dplyr)
library(fdrtool)
library(betareg)
library(splines)
library(Hmisc)
library(ggplot2)
library(reshape2)
library(swfdr)
```

Load the .RData file with the BMI GIANT GWAS meta-analysis data:

```{r,cache=TRUE}
load("BMI_GIANT_GWAS.RData")

tot <- BMI_GIANT_GWAS
```

### Estimate fraction of true null hypotheses in a regression framework

Create the design matrix, using natural cubic splines with 5 degrees of freedom to model `N` and 3 discrete categories for the MAFs:

```{r}
X <- model.matrix(~ splines::ns(N,5) + Freq_MAF_Int_Hapmap, data = tot)[,-1]
dim(X)
head(X)
```

Run code to estimate the fraction of true null hypotheses within a regression framework with the design matrix specified above:

```{r,cache=TRUE}
pi0Est <- lm_pi0(pValues=tot$p, X=X, smooth.df=3)
##caution: this should take about an hour!
```

Save results:

```{r}
fitted0.8 <- pi0Est$pi0.lambda[,pi0Est$lambda==0.8]
fitted0.9 <- pi0Est$pi0.lambda[,round(pi0Est$lambda,2)==0.9]
fitted.final.smooth <- pi0Est$pi0

save(fitted0.8, fitted0.9, fitted.final.smooth, file="BMI_GIANT_GWAS_results.RData")
```

### Session Information

```{r}
devtools::session_info()
```

