---
title: "Generate plot in Figure 4"
author: "Simina Boca, Jeff Leek"
date: "`r Sys.Date()`"
output: rmarkdown::pdf_document
vignette: >
  %\VignetteIndexEntry{swfdr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!--
%% \VignetteEngine{knitr::rmarkdown}
-->

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.path='Figures/')
```

```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown()
```

Load the relevant libraries:

```{r,message=FALSE}
library(ggplot2)
library(reshape2)
library(dplyr)
library(scales)
library(qvalue)
library(xtable)
```

Load the data and the results and add them to the data object:

```{r}
load("BMI_GIANT_GWAS.RData")

tot <- BMI_GIANT_GWAS

load("BMI_GIANT_GWAS_results_logistic.RData")

tot$fitted0.8 <- fitted0.8
tot$fitted0.9 <- fitted0.9
tot$fitted.final.smooth <- fitted.final.smooth

load("BMI_GIANT_GWAS_results_Scott_theoretical.RData")

tot$ScottT <- pi0EstScott

load("BMI_GIANT_GWAS_results_Scott.RData")

tot$ScottE <- pi0EstScott

load("BMI_GIANT_GWAS_bootstrap_CIs_logistic.RData")

dim(tot)
dim(perc.05.95)
head(tot)
head(perc.05.95)
head(rownames(perc.05.95))
identical(as.integer(rownames(perc.05.95)),1:nrow(perc.05.95))

tot$lower.CI <- perc.05.95[,1]
tot$upper.CI <- perc.05.95[,2]


```

Get range of estimates:

```{r}
range(tot$fitted.final.smooth)
range(tot$fitted0.8)
range(tot$fitted0.9)
```

Only use random sample of 50,000 SNPs in plot:

```{r}
##get random sample of 50,000 SNPs
set.seed(310841)
d1 = tot %>% sample_n(5e4)
```

Create long data frame (for ggplot):

```{r}
colnames(d1)[colnames(d1)=="ScottT"] <- "fitted.ScottT"
colnames(d1)[colnames(d1)=="ScottE"] <- "fitted.ScottE"

d3 <- melt(d1,
           id.vars=colnames(d1)[-grep("fitted",colnames(d1))],
           value.name = "pi0",variable.name = "lambda")
d3$lambda <- as.character(d3$lambda)
d3$lambda[d3$lambda=="fitted0.8"] <- "lambda=0.8"
d3$lambda[d3$lambda=="fitted0.9"] <- "lambda=0.9"
d3$lambda[d3$lambda=="fitted.final.smooth"] <- "final smoothed pi0(x)"
d3$lambda[d3$lambda=="fitted.ScottT"] <- "Scott theoretical"
d3$lambda[d3$lambda=="fitted.ScottE"] <- "Scott empirical"

d3$Freq_MAF_Int_Hapmap <- paste("MAF in", d3$Freq_MAF_Int_Hapmap)

head(d3)
```

Function to change the way labels are displayed:

```{r}
display_thousands <- function(l) {
  ##divide by 1000
  ll <- l/1000
  ##add k afterwards
  lll <- paste(ll,"k",sep="")
  
  lll
}
```

Generate Figure 4:

```{r Fig4,cache=TRUE, fig.width=10, fig.height=3}
ggplot(d3, aes(x=N, y=pi0))+
  geom_vline(xintercept=median(tot$N),colour="grey50")+
  geom_ribbon(aes(ymin=lower.CI, ymax=upper.CI, group=Freq_MAF_Int_Hapmap),fill="grey50")+
  geom_line(aes(linetype=lambda)) +
  facet_grid(. ~ Freq_MAF_Int_Hapmap) +
  scale_x_continuous(labels=comma) +
  ylab("Estimated proportion of nulls") +
  scale_linetype_discrete(labels =
                            c(expression(paste("Final smoothed ",
                                               hat(pi)[0](x))),
                              expression(paste(hat(pi)[0], ""^lambda,
                                               (x),
                                               " for ", lambda, "=",
                                               0.8)),
                              expression(paste(hat(pi)[0], ""^lambda,
                                               (x),
                                               " for ", lambda, "=",
                                               0.9)),
                              "Scott empirical",
                              "Scott theoretical")) +
  theme(axis.line = element_line(colour = "black"),
        plot.title = element_text(size = 16, hjust = 0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.key = element_blank(),
        legend.text.align = 0,
        legend.title=element_text(size=13),
        legend.text=element_text(size=12),
        strip.background = element_blank(),
        strip.text.x = element_text(size = 12),
        axis.text.x = element_text(size=10),
        axis.text.y = element_text(size=10))+
        ##axis.line.x = element_line(color="black", size = 0.5), ##this is to show axes - bug in this version of ggplot2
        ##axis.line.y = element_line(color="black", size = 0.5)) +
  guides(linetype=guide_legend(title="Estimate"))

```

Get FDR for our method:

```{r}
##first get BH q-values
qValsBH <- p.adjust(tot$p, method="BH")
##also get Storey q-values
qValsStorey <- qvalue(tot$p)$qvalues
##what is their difference
head(qValsStorey/qValsBH)

qValsReg <- qValsBH*tot$fitted.final.smooth

tot$qValsBH <- qValsBH
tot$qValsStorey <- qValsStorey
tot$qValsReg <- qValsReg

sum(tot$qValsBH < 0.05)
sum(tot$qValsStorey < 0.05)
sum(tot$qValsReg < 0.05)
sum(FDRScott_emp < 0.05)
sum(FDRScott_theo < 0.05)

mean(tot$qValsBH < 0.05)
mean(tot$qValsStorey < 0.05)
mean(tot$qValsReg < 0.05)
mean(FDRScott_emp < 0.05)
mean(FDRScott_theo < 0.05)

set.seed(310841)
d1 = tot %>% sample_n(5e4)

par(mfrow=c(1,2))
plot(d1$qValsReg ~ d1$qValsStorey, xlim=c(0,1), ylim=c(0,1))
abline(a=0,b=1, col="red", lwd=2)
plot(d1$qValsReg ~ d1$qValsBH, xlim=c(0,1), ylim=c(0,1))
abline(a=0,b=1, col="red", lwd=2)

```

Make nice table for summary results:
```{r}
xtable(data.frame(BL = sum(tot$qValsReg <= 0.05),
                  Scott_T = sum(FDRScott_theo <= 0.05),
                  Scott_E = sum(FDRScott_emp <= 0.05),
                  Storey = sum(tot$qValsStorey <= 0.05),
                  BH = sum(tot$qValsBH <= 0.05)))
```

How many are in common between all the methods?
```{r, results='asis'}
length(Reduce(intersect, list(which(tot$qValsReg <= 0.05), 
                              which(FDRScott_theo <= 0.05), 
                              which(FDRScott_emp <= 0.05),
                              which(tot$qValsStorey <= 0.05), 
                              which(tot$qValsBH <= 0.05))))

length(Reduce(intersect, list(which(tot$qValsReg <= 0.05), 
                              which(tot$qValsStorey <= 0.05), 
                              which(tot$qValsBH <= 0.05))))

length(Reduce(intersect, list(which(tot$qValsStorey <= 0.05), 
                              which(tot$qValsReg <= 0.05))))

length(Reduce(intersect, list(which(FDRScott_theo <= 0.05), 
                              which(tot$qValsReg <= 0.05))))

length(Reduce(intersect, list(which(FDRScott_theo <= 0.05), 
                              which(FDRScott_emp <= 0.05),
                              which(tot$qValsBH <= 0.05))))

length(Reduce(intersect, list(which(FDRScott_emp <= 0.05),
                              which(tot$qValsBH <= 0.05))))

length(Reduce(intersect, list(which(FDRScott_theo <= 0.05), 
                              which(tot$qValsBH <= 0.05))))
```


### Session Information

```{r}
devtools::session_info()
```

